---
title: "How Embryos Acclimate to Temperature through Epigenetic Regulation"
subtitle: "Scripts"
author: 
  - "Thomas O’Leary, Emily Mikucki, Sumaetee Tangwancharoen, Sara Helms Cahan, Seth Frietze, and Brent L. Lockwood"
output:
  rmarkdown::html_document:
    theme: lumen
    number_sections: true
    toc: true
    toc_float: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# HEATER {-}

## Goals {-}

HEATER Project:

# **Embryonic thermal acclimation**

## Egg hatching success

```{r, file = "src/00_pheno/cs_pheno.R"}
```

# **Nuclei isolation**

## Nuclei quality

## Nuclei counts

The data is loaded and then the species are grouped.

```{r, file = "src/00_pheno/cs_pheno_fig.R"}
```

# **Cell Ranger ARC**

## `mkref`

```{r, file = "src/02_cellranger-arc/00_mkref.sh"}
```

## `count`

```{r, file = "src/02_cellranger-arc/01_count.sh"}
```

## `aggr`

```{r, file = "src/02_cellranger-arc/02_aggr.sh"}
```

# **Seurat**

## Create seurat object

```{r, file = "src/03_seurat/01_create_seurat_object.R"}
```

## Quality control and filtering

```{r, file = "src/03_seurat/02_quality_control_filtering.R"}
```

## Clustering

```{r, file = "src/03_seurat/03_cluster.R"}
```

## Cluster Plots

```{r, file = "src/03_seurat/04_cluster_plots.R"}
```

## Cluster Markers

```{r, file = "src/03_seurat/05_cluster_markers.R"}
```

## Differentially expressed genes

```{r, file = "src/03_seurat/06_degs.R"}
```

## Feature plots

```{r, file = "src/03_seurat/08_feature_plots.R"}
```

# Session Info {-}

```{r}
sessionInfo()
```
